---
title: "lab_06_answers"
author: "Cassie Peo"
date: "2024-02-29"
output: html_document
---

## Load libraries and establish settings
**Task**: Load janitor and the tidyverse
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse,rvest,janitor

library(tidyverse)

library(janitor)

library(rvest)

```


Q1. We are using NCAA Demographics Data showing race, gender for athletic directors and student athletes. ncaa_athletic_dirs.csv

Background on the data:
https://www.ncaa.org/sports/2018/12/13/ncaa-demographics-database.aspx?id=1729

Load the data into a dataframe called ads1. 
Create another dataframe where this data is filtered for 2021 only.
Using ggplot, create a basic column chart showing athletic directors by race and sex. 
Write a headline and caption.

**A1**

```{r}

ads1 <- read_csv("ncaa_athletic_dirs.csv") %>% 
  as.data.frame()

ads1_2021 <- ads1 %>%
  filter(year == 2021)

```

```{r}
ggplot(ads1_2021, aes(race_sex, count, fill = count)) +
  geom_col(position = "dodge") +
  geom_text(aes(label=count), vjust = -.5, size = 2.5) +
  labs(
    title="The Majority of NCAA Athletic Directors Were White Men in 2021",
    x = "race_sex",
    y = "count",
    caption = "There were 705 white male athletic directors in 2021. 
    source: NCAA Demographics Data showing race, gender for athletic directors")

```



Q2. Now create a second chart using ads1 (ncaa_athletic_dirs) into a chart of columns with years on the x axis.
Write a headline and caption.

**A2**

```{r}
 
ggplot(ads1, aes(year, count, fill = race_sex)) +
  geom_col(position = "dodge") +
  labs(
    title="The Race and Sex of the NCAA Athletic Directors 2012-2021",
    x = "year",
    y = "count",
    caption = "From 2012 through 2021, the majority of NCAA Athletic Directors were white males. 
    source:NCAA Demographics Data showing race, gender for athletic directors")


```


Q3. We will now visualize the data over the years for white men, black men and white women and black women athletic directors using this data: Sports_Race_ADs.csv

Remember to clean the names. Import the data with the appropriate tab and it beginning with the header row.

For example:
dataframe <- read_csv('my_spreadsheet.csv', skip =1)
Skip =1 allows the data to be imported beginning with the second row. This is done with some spreadsheets that have headlines and other junk in the first row.


Load the data, create a line chart showing athletic directors by race & sex. 
Designate one line for each and a separate color: ie male_white would be blue, female_other would be pink, etc. 

**A3.**

```{r}

Sports_Race_ADs <- read_csv('Sports_Race_ADs.csv', skip =1)

#This cleans column names
Sports_Race_ADs <- janitor::clean_names(Sports_Race_ADs)

ggplot(Sports_Race_ADs, aes(x=year, y=total)) +
  geom_line(aes(y=male_white), color = "blue") +
  geom_line(aes(y=male_black), color = "red") +
  geom_line(aes(y=male_other), color = "green") +
  geom_line(aes(y=female_white), color = "pink") +
  geom_line(aes(y=female_black), color = "purple") +
  geom_line(aes(y=female_other), color = "yellow")

```