---
title: "r_visualizing_reporting"
author: "Cassie Peo"
date: "2024-02-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Pre-Lab Question 1
Which year had the most arrests? The arrest code is ARR.
Create a new dataframe that filters for arrests and totals them by year. Call this arrest.

```{r}

arrest <- SF %>%
  filter(disposition == "ARR")

arrest <- arrest %>% 
  select(year, disposition) %>% 
  group_by(year) %>% 
  count(disposition)

# 2018 had the most arrests with 159 arrests. 

```


Then create a bar chart that colors the bars by arrest total. Include a headline and caption
```{r}

arrest %>%
  ggplot(aes(x=year, y=n, fill=n)) +
  coord_flip() + 
  theme_economist() +
  geom_col()+
  geom_text(aes(label=n), hjust = -.1, size = 2.5) +
    labs(
    title="311 Calls Resulted in Arrests in San Francisco, 2016-2019",
    x = "year",
    y = "total arrests",
    caption = "2018 had the most arrests with 159 arrests.
    Source: SF PD Service Call Data, 2016-2019")


```

# Pre-Lab Question 2
What are the most common resolutions for 311 calls?
Build a table called all_dispo that counts all of the dispositions
Filter to the top 12 most common dispositions.

And here are the disposition codes, in case you're curious:
https://github.com/profrobwells/Guest_Lectures/blob/5c2d631aec862cd728ebf49d71143dbfcade3567/Intro_to_R_2022/Radio_Codes_2016.xlsx

```{r}

all_dispo <- SF %>% 
  select(year, disposition) %>% 
  count(disposition)

top12_all_dispo <- all_dispo %>%
  slice_max(n, n=12)

# Code HAN, or "Handled" is the most common resolution for 311 calls, with 51,064 handled resolutions. The second most common resolution is ADV or "Advised," with 49,823 advised resolutions. The third most common resolution is GOA or "Gone on Arrival," with 31,951 resolutions. The top 12 resolution codes in order from most to least common are HAN, ADV, GOA, CIT, UTL, Not recorded, ND, CAN, ADM, REP, NOM, and ARR. 

```

Create a bar chart of all_dispo and include a headline and caption


```{r}

top12_all_dispo %>%
  ggplot(aes(x=disposition, y=n, fill=n)) +
  coord_flip() + 
  theme_economist() +
  geom_col()+
  theme(legend.text = element_text(size=5)) +
  geom_text(aes(label=n), hjust = -.1, size = 2) +
    labs(
    title="The Top 12 Most Common Resolutions for 311 Calls",
    x = "disposition",
    y = "number of instances",
    caption = "Code HAN, or Handled is the most common resolution for 311 calls, with 51,064 handled resolutions.  
    source: SF PD Service Call Data, 2016-2019")



```

**NERD ZONE**
# Pre-Lab Question 3
Take the chart above, sort the bars in descending order
```{r}

top12_all_dispo %>%
  ggplot(aes(x=reorder(disposition, n), y=n, fill=n)) +
  coord_flip() + 
  theme_economist() +
  geom_col()+
  theme(legend.text = element_text(size=5)) +
  geom_text(aes(label=n), hjust = -.1, size = 2) +
    labs(
    title="The Top 12 Most Common Resolutions for 311 Calls",
    x = "disposition",
    y = "number of instances",
    caption = "Code HAN, or Handled is the most common resolution for 311 calls, with 51,064 handled resolutions.  
    source: SF PD Service Call Data, 2016-2019")


```