---
title: "A1_draft_notebook"
author: "Cassie Peo"
format: html
---

For this assignment, I would like to examine the demographics of Baltimore’s top and bottom 10% wealthiest neighborhoods. In this course, we have already worked with Maryland and Baltimore income and race data, and I am interested in seeing how the demographics, like race and gender, differ in these neighborhoods. I want to use the datasets ‘baltcity_income_clean’ and ‘baltcity_race’ that we have already worked with to answer the question “how do the demographics of Baltimore’s top 10% wealthiest neighborhoods compare to the demographics of Baltimore’s bottom 10% wealthiest neighborhoods?” Then, I found a dataset about female-headed households in Baltimore and am interested in exploring how the numbers of female-headed households differ in the wealthiest and poorest neighborhoods in Baltimore to answer questions like “are there more female-headed households in lower income areas?” and “what is the percentage of female-headed households in Baltimore’s top and bottom 10% wealthiest neighborhoods?”

Topic: 
How do the demographics of Baltimore’s top 10% wealthiest neighborhoods compare to the demographics of Baltimore’s bottom 10% wealthiest neighborhoods?

Load in libraries:
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse,rvest,janitor

library(tidyverse)

library(janitor)

library(rvest)

library(dplyr)

library(stringr)
```


Load in data:
```{r}
baltcity_income <- read_csv("baltcity_income_clean.csv") %>% 
  as.data.frame()

#This cleans column names
baltcity_income <- janitor::clean_names(baltcity_income)

baltcity_race <- read_csv("baltcity_race_8_13.csv") %>% 
  as.data.frame()

#This cleans column names
baltcity_race <- janitor::clean_names(baltcity_race)

pct_female_headed_households <- read_csv("Percent_of_Female-Headed_Households_with_Children_Under_18.csv") %>% 
  as.data.frame()

#This cleans column names
pct_female_headed_households <- janitor::clean_names(pct_female_headed_households)


updated_baltcity_race <- baltcity_income %>%
  inner_join(baltcity_race, by="geoid")
```

Questions:

1. What is the percentage of non-white residents in Baltimore’s top 10% wealthiest neighborhoods?

```{r}
updated_baltcity_race %>%
  slice_max(x2020, n = 20) %>%
  mutate(x2020_nonwhite=((x2020_total-x2020_white)/x2020_total)) %>%
  select(x2020_nonwhite, neighborhood, x2020) %>%
arrange(desc(x2020))

# As the code above shows, the percentage of non-white residents in Baltimore’s wealthiest neighborhood, North Baltimore/Guilford/Homeland, is 38.1%. Baltimore's second wealthiest neighborhood, North Baltimore/Guilford/Homeland, has 23.42% non-white residents. The third wealthiest neighborhood, Greater Roland Park/Poplar Hill, has a percentage of 33.24% non-white residents. The fourth wealthiest neighborhood, South Baltimore, has 11.51% non-white residents, and the fifth wealthiest neighborhood, Canton, has 13.15% non-white residents.
```


2. What is the percentage of non-white residents in Baltimore’s bottom 10% wealthiest neighborhoods? 

```{r}
updated_baltcity_race %>%
  slice_min(x2020, n = 20) %>%
  mutate(x2020_nonwhite=((x2020_total-x2020_white)/x2020_total)) %>%
  select(x2020_nonwhite, neighborhood, x2020) %>%
arrange(x2020)

# As the code above shows, the percentage of non-white residents in Baltimore’s poorest neighborhood, Upton/Druid Heights, is 90.55%. Sandtown-Winchester/Harlem Park the second poorest neighborhood's percentage of non-white residents is 96.22%, and the third poorest neighborhood, Southwest Baltimore, has a percentage of 69.8% non-white residents. 
```


3. What is the average income difference between these two sets of neighborhoods?

```{r}
top_10 <- updated_baltcity_race %>%
  slice_max(x2020, n = 20)
top_10_mean <- mean(top_10$x2020)

bottom_10 <- updated_baltcity_race %>%
  slice_min(x2020, n = 20)
bottom_10_mean <- mean(bottom_10$x2020)

top_10_mean - bottom_10_mean
(top_10_mean - bottom_10_mean) / bottom_10_mean

# As shown by the code above, the average income of the top 10% is $109496.40 more than the average income of the bottom 10%. The top 10% on average earns 5.26 times more than the bottom 10% on average.  
```

4. What is the percentage of female-headed households in Baltimore’s top and bottom 10% wealthiest neighborhoods?

```{r}
pct_female_headed_households <- pct_female_headed_households %>%
  rename(neighborhood = csa2020)

female_cmbd <- updated_baltcity_race %>%
  inner_join(pct_female_headed_households, by="neighborhood")

top_pcts <- distinct(top_10 %>%
           select(neighborhood)) %>%
  inner_join(female_cmbd, by = "neighborhood") %>%
  select(femhhs21)

mean(top_pcts$femhhs21)

bottom_pcts <- distinct(bottom_10 %>%
           select(neighborhood)) %>%
  inner_join(female_cmbd, by = "neighborhood") %>%
  select(femhhs21)

mean(bottom_pcts$femhhs21)

# As the code above shows, the percentage of female-headed households in Baltimore’s top 10% wealthiest neighborhoods is 22.06%. The percentage of female-headed households in Baltimore’s bottom 10% wealthiest neighborhoods is 66.82%. 

```

