---
title: 'Assignment 1, Step #3'
author: "Cassie Peo"
date: "2024-03-13"
output: html_document
---

For this assignment, I examined the demographics of Baltimore’s top and bottom 10% wealthiest neighborhoods. In this course, we have already worked with Maryland and Baltimore income and race data, and I was interested in seeing how the demographics, like race and gender, differ in these neighborhoods. I used the datasets ‘baltcity_income_clean’ and ‘baltcity_race’ that we have already worked with to answer the question “how do the demographics of Baltimore’s top 10% wealthiest neighborhoods compare to the demographics of Baltimore’s bottom 10% wealthiest neighborhoods?” Then, I used a dataset about female-headed households in Baltimore to explore how the numbers of female-headed households differ in the wealthiest and poorest neighborhoods in Baltimore. I aimed to answer questions like “are there more female-headed households in lower income areas?” and “what is the percentage of female-headed households in Baltimore’s top and bottom 10% wealthiest neighborhoods?”
Using these datasets, I found that the percentage of non-white residents in Baltimore’s wealthiest neighborhood, North Baltimore/Guilford/Homeland, is 38.1%. Baltimore's second wealthiest neighborhood, North Baltimore/Guilford/Homeland, has 23.42% non-white residents. The third wealthiest neighborhood, Greater Roland Park/Poplar Hill, has a percentage of 33.24% non-white residents. 
The percentage of non-white residents in Baltimore’s poorest neighborhood, Upton/Druid Heights, is 90.55%. Sandtown-Winchester/Harlem Park the second poorest neighborhood's percentage of non-white residents is 96.22%, and the third poorest neighborhood, Southwest Baltimore, has a percentage of 69.8% non-white residents.
In the income data, the most recent year was 2020, but the female-headed household data had data for years 2010, 2015, 2016, 2017, 2018, 2019, and 2021. There was no data for 2020, so I used the data from 2021 for the percentage of female-headed households calculations and graphics. 
The percentage of female-headed households in Baltimore’s top 10% wealthiest neighborhoods is 22.06%. The percentage of female-headed households in Baltimore’s bottom 10% wealthiest neighborhoods is 66.82%. 
Harbor East/Little Italy had the highest percentage of female-headed households in the top 10% wealthiest neighborhoods in Baltimore with 71.4% female-headed households. Highlandtown had the second highest, with 35.9%, and Patterson Park North & East had the third highest with 34.2% female-headed households. 
Upton/Druid Heights had the highest percentage of female-headed households in the bottom 10% wealthiest neighborhoods in Baltimore with 86.6% female-headed households. Upton/Druid Heights also had the highest percentage of non-white residents -- 90.55%. Cherry Hill had the second highest percentage of female-headed households with 81% and Westport/Mount Winans/Lakeland had the third highest with 73.1%. 
It is very interesting to me that most of the neighborhoods in Baltimore’s top 10% wealthiest neighborhoods had less than 36% female-headed households while Harbor East/Little Italy had the highest percentage of female-headed households with 71.4%. For a story, I would like to further investigate why this is and what factors make Harbor East/Little Italy different from the rest of the top 10% wealthiest neighborhoods. I would want to talk to people in Harbor East/Little Italy, as well as the rest of the top 10% wealthiest neighborhoods. 

Tweet 1: In the bottom 10% wealthiest neighborhoods in Baltimore, Upton/Druid Heights had the highest percentage of female-headed households with 86.6% female-headed households, and the highest percentage of non-white residents at 90.55%.

Tweet 2: The top 10% wealthiest neighborhoods in Baltimore, on average, earn 5.26 times more than the bottom 10% on average. The average income of the top 10% is $109496.40 more than the average income of the bottom 10%. 

Topic: 
How do the demographics of Baltimore’s top 10% wealthiest neighborhoods compare to the demographics of Baltimore’s bottom 10% wealthiest neighborhoods?

Load in libraries:
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse,rvest,janitor

library(tidyverse)

library(janitor)

library(rvest)

library(dplyr)

library(stringr)
```

Load in data:
```{r}
baltcity_income <- read_csv("baltcity_income_clean.csv") %>% 
  as.data.frame()

#This cleans column names
baltcity_income <- janitor::clean_names(baltcity_income)

baltcity_race <- read_csv("baltcity_race_8_13.csv") %>% 
  as.data.frame()

#This cleans column names
baltcity_race <- janitor::clean_names(baltcity_race)

pct_female_headed_households <- read_csv("Percent_of_Female-Headed_Households_with_Children_Under_18.csv") %>% 
  as.data.frame()

#This cleans column names
pct_female_headed_households <- janitor::clean_names(pct_female_headed_households)


updated_baltcity_race <- baltcity_income %>%
  inner_join(baltcity_race, by="geoid")
```

Questions:

1. What is the percentage of non-white residents in Baltimore’s top 10% wealthiest neighborhoods?

```{r}
updated_baltcity_race %>%
  slice_max(x2020, n = 20) %>%
  mutate(x2020_nonwhite=((x2020_total-x2020_white)/x2020_total)) %>%
  select(x2020_nonwhite, neighborhood, x2020) %>%
arrange(desc(x2020))

top20_baltcity <- updated_baltcity_race %>%
  slice_max(x2020, n = 20) %>%
  mutate(x2020_nonwhite=((x2020_total-x2020_white)/x2020_total)) %>%
  select(x2020_nonwhite, neighborhood, x2020, geoid) %>%
arrange(desc(x2020))

# As the code above shows, the percentage of non-white residents in Baltimore’s wealthiest neighborhood, North Baltimore/Guilford/Homeland, is 38.1%. Baltimore's second wealthiest neighborhood, North Baltimore/Guilford/Homeland, has 23.42% non-white residents. The third wealthiest neighborhood, Greater Roland Park/Poplar Hill, has a percentage of 33.24% non-white residents. The fourth wealthiest neighborhood, South Baltimore, has 11.51% non-white residents, and the fifth wealthiest neighborhood, Canton, has 13.15% non-white residents.
```

1a. Bar chart of the top 10% wealthiest neighborhoods in Baltimore.

```{r}

top5_baltcity <- updated_baltcity_race %>%
  slice_max(x2020, n = 5) %>%
  mutate(x2020_nonwhite=((x2020_total-x2020_white)/x2020_total)) %>%
  select(x2020_nonwhite, neighborhood, x2020, geoid) %>%
arrange(desc(x2020))

top5_baltcity$neighborhood <- paste(top5_baltcity$neighborhood, top5_baltcity$geoid, sep="-")

ggplot(top5_baltcity, aes(neighborhood, x2020, fill = x2020)) +
  coord_flip() +
  geom_col(position = "dodge") +
  geom_text(aes(label=x2020), vjust = -.5, size = 2.5) +
  labs(
    title="The Top 10% Wealthiest Neighborhoods in Baltimore in 2020",
    x = "Neighborhood",
    y = "Median Income",
    caption = "North Baltimore/Guilford/Homeland, North Baltimore/Guilford/Homeland, 
    and Greater Roland Park/Poplar Hill are the top three wealthiest neighborhoods.  
    source: baltcity_income_clean.csv")

# As the code above shows, Baltimore’s wealthiest neighborhood is North Baltimore/Guilford/Homeland. Baltimore's second wealthiest neighborhood is North Baltimore/Guilford/Homeland. The third wealthiest neighborhood is Greater Roland Park/Poplar Hill. The fourth wealthiest neighborhood is South Baltimore and the fifth wealthiest neighborhood is Canton.

```

2. What is the percentage of non-white residents in Baltimore’s bottom 10% wealthiest neighborhoods? 

```{r}
updated_baltcity_race %>%
  slice_min(x2020, n = 20) %>%
  mutate(x2020_nonwhite=((x2020_total-x2020_white)/x2020_total)) %>%
  select(x2020_nonwhite, neighborhood, x2020) %>%
arrange(x2020)

# As the code above shows, the percentage of non-white residents in Baltimore’s poorest neighborhood, Upton/Druid Heights, is 90.55%. Sandtown-Winchester/Harlem Park the second poorest neighborhood's percentage of non-white residents is 96.22%, and the third poorest neighborhood, Southwest Baltimore, has a percentage of 69.8% non-white residents. 
```

3. What is the average income difference between these two sets of neighborhoods?

```{r}
top_10 <- updated_baltcity_race %>%
  slice_max(x2020, n = 20)
top_10_mean <- mean(top_10$x2020)

bottom_10 <- updated_baltcity_race %>%
  slice_min(x2020, n = 20)
bottom_10_mean <- mean(bottom_10$x2020)

top_10_mean - bottom_10_mean
(top_10_mean - bottom_10_mean) / bottom_10_mean

# As shown by the code above, the average income of the top 10% is $109496.40 more than the average income of the bottom 10%. The top 10% on average earns 5.26 times more than the bottom 10% on average.  
```


4. What is the percentage of female-headed households in Baltimore’s top and bottom 10% wealthiest neighborhoods?

```{r}
pct_female_headed_households <- pct_female_headed_households %>%
  rename(neighborhood = csa2020)

female_cmbd <- updated_baltcity_race %>%
  inner_join(pct_female_headed_households, by="neighborhood")

top_pcts <- distinct(top_10 %>%
           select(neighborhood)) %>%
  inner_join(female_cmbd, by = "neighborhood") %>%
  select(femhhs21, neighborhood)

mean(top_pcts$femhhs21)

bottom_pcts <- distinct(bottom_10 %>%
           select(neighborhood)) %>%
  inner_join(female_cmbd, by = "neighborhood") %>%
  select(femhhs21, neighborhood)

mean(bottom_pcts$femhhs21)

# As the code above shows, the percentage of female-headed households in Baltimore’s top 10% wealthiest neighborhoods is 22.06%. The percentage of female-headed households in Baltimore’s bottom 10% wealthiest neighborhoods is 66.82%. 

```

4a. Bar chart of the percentage of female-headed households in top 10% wealthiest neighborhoods in Baltimore.

```{r}
ggplot(top_pcts, aes(reorder(neighborhood, femhhs21), femhhs21, fill = femhhs21)) +
  coord_flip() +
  geom_col(position = "dodge") +
  geom_text(aes(label=femhhs21), vjust = -.5, size = 2.5) +
  labs(
    title="The Percentage of Female-headed Households in the 
    Top 10% Wealthiest Neighborhoods in Baltimore in 2021",
    x = "Neighborhood",
    y = "Percentage of Female-headed Households",
    caption = "Harbor East/Little Italy had the highest percentage of female-headed households 
    in the top 10% wealthiest neighborhoods in Baltimore with 71.4% female-headed households.  
    source: baltcity_income_clean.csv and Percent_of_Female-Headed_Households_with_Children_Under_18.csv")

# As the code above shows, Harbor East/Little Italy had the highest percentage of female-headed households in the top 10% wealthiest neighborhoods in Baltimore with 71.4% female-headed households. Highlandtown had the second highest, with 35.9%, and Patterson Park North & East had the third highest with 34.2% female-headed households. 
```

4b. Bar chart of the percentage of female-headed households in bottom 10% wealthiest neighborhoods in Baltimore.

```{r}
ggplot(bottom_pcts, aes(reorder(neighborhood, femhhs21), femhhs21, fill = femhhs21)) +
  coord_flip() +
  geom_col(position = "dodge") +
  geom_text(aes(label=femhhs21), vjust = -.5, size = 2.5) +
  labs(
    title="The Percentage of Female-headed Households in the 
    Bottom 10% Wealthiest Neighborhoods in Baltimore in 2021",
    x = "Neighborhood",
    y = "Percentage of Female-headed Households",
    caption = "Upton/Druid Heights had the highest percentage of female-headed households in the 
    bottom 10% wealthiest neighborhoods in Baltimore with 86.6% female-headed households.  
    source: baltcity_income_clean.csv and Percent_of_Female-Headed_Households_with_Children_Under_18.csv")

# As the code above shows, Upton/Druid Heights had the highest percentage of female-headed households in the bottom 10% wealthiest neighborhoods in Baltimore with 86.6% female-headed households. Cherry Hill had the second highest percentage of female-headed households with 81% and Westport/Mount Winans/Lakeland had the third highest with 73.1%.

```
